---
title: "Project 1 Markdown MV"
author: "Manuel V."
date: "2023-04-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions
The World Health Organization has recently employed a new data science initiative, *CSIT-165*, that uses data science to characterize pandemic diseases. 
*CSIT-165* disseminates data driven analyses to global decision makers.

*CSIT-165* is a conglomerate comprised of two fabricated entities: *Global Health Union (GHU)* and *Private Diagnostic Laboratories (PDL)*. 
Your and your partner's role is to play a data scientist from one of these two entities.


## Data information
Data for 2019 Novel Coronavirus is operated by the John Hopkins University Center for Systems Science and Engineering (JHU CSSE). 
Data includes daily time series CSV summary tables, including confirmations, recoveries, and deaths. Country/region are countries/regions hat conform to World Health Organization (WHO). 
Lat and Long refer to coordinates references for the user. 
Date fields are stored in MM/DD/YYYY format.

## Project Objectives

### Objective 1

```{r ob1}
# Get csv files from CSSEGISandData/COVID-19 database

covid_confirmations <- read.csv("https://raw.githubusercontent.com/DivaMC/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv", stringsAsFactors=FALSE)

covid_deaths <- read.csv("https://raw.githubusercontent.com/DivaMC/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv", stringsAsFactors=FALSE)

covid_recoveries <- read.csv("https://raw.githubusercontent.com/DivaMC/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_recovered_global.csv", stringsAsFactors=FALSE)


# Find greatest number of deaths on first day

for(i in covid_deaths$X1.22.20) {
  if(i > 0) {
    print(i)
  }
}

max(covid_deaths$X1.22.20)
first_city_deaths <- covid_deaths$Province.State[which.max(covid_deaths$X1.22.20)]


# Find greatest number of confirmed cases on first day

max(covid_confirmations$X1.22.20)
first_city_confirmation <- covid_confirmations$Province.State[which.max(covid_confirmations$X1.22.20)]


# Find and print city of origin

if (first_city_deaths == first_city_confirmation) {
  covid_city_origin <- first_city_confirmation
}

print(covid_city_origin)



```

### Objective 2

```{r ob2}
## Most recent area to have a confirmed case
## Use a for loop, if statement, and subsets.

recent_confirmations <- c()

for (i in 1:nrow(covid_confirmations)) {
  if (covid_confirmations$X3.9.23[i] > covid_confirmations$X3.8.23[i]) {
    recent_confirmations <- c(recent_confirmations,covid_confirmations$Country.Region[i])
   }
}

print(recent_confirmations)

# R Solution
# recent_confirmations <- paste(covid_confirmations$Province.State[which(covid_confirmations$X3.9.23 > covid_confirmations$X3.8.23)], 
#                              covid_confirmations$Country.Region[which(covid_confirmations$X3.9.23 > covid_confirmations$X3.8.23)])


```

### Objective 3

```{r ob3}
library ("geosphere")

# How far away are the areas from objective 2 from where the first confirmed case(s) occurred?

first_city_confirmation_index <- which.max(covid_confirmations$X1.22.20)

area_df <- data.frame(Long = covid_confirmations$Long[which(covid_confirmations$X3.9.23 > covid_confirmations$X3.8.23)]
                      ,Lat = covid_confirmations$Lat[which(covid_confirmations$X3.9.23 > covid_confirmations$X3.8.23)])

area_df_matrix <- as.matrix(area_df)

# First City Coordinates

First_City_Coords <- data.frame(Long = covid_confirmations$Long[first_city_confirmation_index]
                           ,Lat = covid_confirmations$Lat[first_city_confirmation_index]) 

First_City_Coords_matrix <- as.matrix(First_City_Coords)

# Find distance between cities
# Use function distm
# distm(x, y)

dist <- distm(area_df_matrix, First_City_Coords_matrix)

# Convert from meters to miles

dist_miles <- round(dist/1609.34, 2)


# Please print the following:
# {recent region} is {distance in miles} away from {origin city, origin country}

distance_from_origin <- data.frame(City = covid_confirmations$Province.State[which(covid_confirmations$X3.9.23 > covid_confirmations$X3.8.23)]
                                   ,Country = covid_confirmations$Country.Region[which(covid_confirmations$X3.9.23 > covid_confirmations$X3.8.23)]
                                   ,Distance = dist_miles)

for(i in 1:length(distance_from_origin$City)) {
  print(paste(distance_from_origin$City[i], distance_from_origin$Country[i], " is ", 
              distance_from_origin$Distance[i], "miles away from", 
              covid_confirmations$Province.State[first_city_confirmation_index], 
              covid_confirmations$Country.Region[first_city_confirmation_index]))
}

```

### Objective 5
```{r ob5}
# Load the CSV files for COVID-19 data
confirmed_cases <- read.csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv", stringsAsFactors = FALSE)
deaths <- read.csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv", stringsAsFactors = FALSE)

# Extract the country names from the data frames
countries <- unique(confirmed_cases$Country.Region)

# Initialize empty lists for the sums of confirmed cases and deaths
confirmed_sums <- list()
death_sums <- list()

# Calculate the sums of confirmed cases and deaths for each country
for (country in countries) {
  confirmed_sums[[country]] <- sum(confirmed_cases[confirmed_cases$Country.Region == country, -(1:4)])
  death_sums[[country]] <- sum(deaths[deaths$Country.Region == country, -(1:4)])
}

# Convert the sums into data frames
confirmed_df <- data.frame(Country = names(confirmed_sums), Confirmed = unlist(confirmed_sums))
death_df <- data.frame(Country = names(death_sums), Deaths = unlist(death_sums))

# Sort the data frames by the values of confirmed cases and deaths, respectively
confirmed_df <- confirmed_df[order(-confirmed_df$Confirmed),]
death_df <- death_df[order(-death_df$Deaths),]

# Select the top 5 countries with the most confirmed cases and deaths
top_5_confirmed <- head(confirmed_df, 5)
top_5_deaths <- head(death_df, 5)

# Convert the data frames into tables using the kable function
library(knitr)
kable(top_5_confirmed)
kable(top_5_deaths)


```
