---
title: "Project 1 Markdown MV"
author: "Manuel V."
date: "2023-04-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Final GitHub Link https://github.com/Chato1969/Project_1
Collaboartion GitHub https://github.com/DivaMC/Project_1

## Instructions
The World Health Organization has recently employed a new data science initiative, *CSIT-165*, that uses data science to characterize pandemic diseases. 
*CSIT-165* disseminates data driven analyses to global decision makers.

*CSIT-165* is a conglomerate comprised of two fabricated entities: *Global Health Union (GHU)* and *Private Diagnostic Laboratories (PDL)*. 
Your and your partner's role is to play a data scientist from one of these two entities.


## Data information
Data for 2019 Novel Coronavirus is operated by the John Hopkins University Center for Systems Science and Engineering (JHU CSSE). 
Data includes daily time series CSV summary tables, including confirmations, recoveries, and deaths. Country/region are countries/regions hat conform to World Health Organization (WHO). 
Lat and Long refer to coordinates references for the user. 
Date fields are stored in MM/DD/YYYY format.

## Project Objectives

### Objective 1

```{r ob1}
# Get csv files from CSSEGISandData/COVID-19 database

covid_confirmations <- read.csv("https://raw.githubusercontent.com/DivaMC/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv", stringsAsFactors=FALSE)

covid_deaths <- read.csv("https://raw.githubusercontent.com/DivaMC/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv", stringsAsFactors=FALSE)

covid_recoveries <- read.csv("https://raw.githubusercontent.com/DivaMC/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_recovered_global.csv", stringsAsFactors=FALSE)


# Find greatest number of deaths on first day

for(i in covid_deaths$X1.22.20) {
  if(i > 0) {
    print(i)
  }
}

max(covid_deaths$X1.22.20)
first_city_deaths <- covid_deaths$Province.State[which.max(covid_deaths$X1.22.20)]


# Find greatest number of confirmed cases on first day

max(covid_confirmations$X1.22.20)
first_city_confirmation <- covid_confirmations$Province.State[which.max(covid_confirmations$X1.22.20)]


# Find and print city of origin

if (first_city_deaths == first_city_confirmation) {
  covid_city_origin <- first_city_confirmation
}

print(covid_city_origin)



```

### Objective 2

```{r ob2}
## Most recent area to have a confirmed case
## Use a for loop, if statement, and subsets.

recent_confirmations <- c()

for (i in 1:nrow(covid_confirmations)) {
  if (covid_confirmations$X3.9.23[i] > covid_confirmations$X3.8.23[i]) {
    recent_confirmations <- c(recent_confirmations,covid_confirmations$Country.Region[i])
   }
}

print(recent_confirmations)

# R Solution
# recent_confirmations <- paste(covid_confirmations$Province.State[which(covid_confirmations$X3.9.23 > covid_confirmations$X3.8.23)], 
#                              covid_confirmations$Country.Region[which(covid_confirmations$X3.9.23 > covid_confirmations$X3.8.23)])


```

### Objective 3

```{r ob3}
library ("geosphere")

# How far away are the areas from objective 2 from where the first confirmed case(s) occurred?

first_city_confirmation_index <- which.max(covid_confirmations$X1.22.20)

area_df <- data.frame(Long = covid_confirmations$Long[which(covid_confirmations$X3.9.23 > covid_confirmations$X3.8.23)]
                      ,Lat = covid_confirmations$Lat[which(covid_confirmations$X3.9.23 > covid_confirmations$X3.8.23)])

area_df_matrix <- as.matrix(area_df)

# First City Coordinates

First_City_Coords <- data.frame(Long = covid_confirmations$Long[first_city_confirmation_index]
                           ,Lat = covid_confirmations$Lat[first_city_confirmation_index]) 

First_City_Coords_matrix <- as.matrix(First_City_Coords)

# Find distance between cities
# Use function distm
# distm(x, y)

dist <- distm(area_df_matrix, First_City_Coords_matrix)

# Convert from meters to miles

dist_miles <- round(dist/1609.34, 2)


# Please print the following:
# {recent region} is {distance in miles} away from {origin city, origin country}

distance_from_origin <- data.frame(City = covid_confirmations$Province.State[which(covid_confirmations$X3.9.23 > covid_confirmations$X3.8.23)]
                                   ,Country = covid_confirmations$Country.Region[which(covid_confirmations$X3.9.23 > covid_confirmations$X3.8.23)]
                                   ,Distance = dist_miles)

for(i in 1:length(distance_from_origin$City)) {
  print(paste(distance_from_origin$City[i], distance_from_origin$Country[i], " is ", 
              distance_from_origin$Distance[i], "miles away from", 
              covid_confirmations$Province.State[first_city_confirmation_index], 
              covid_confirmations$Country.Region[first_city_confirmation_index]))
}

```

### Objective 4
```{r ob4}
# Import the data for global confirmed cases
confirmed_url <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv"
confirmed_data <- read.csv(confirmed_url, header = TRUE, stringsAsFactors = FALSE)

# Import the data for global deaths
deaths_url <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv"
deaths_data <- read.csv(deaths_url, header = TRUE, stringsAsFactors = FALSE)

# Remove cruise ships from the confirmed cases data
confirmed_data <- confirmed_data[!(confirmed_data$Lat == 0 | is.na(confirmed_data$Lat)), ]
confirmed_data <- confirmed_data[!(confirmed_data$Long == 0 | is.na(confirmed_data$Long)), ]

# Remove cruise ships from the deaths data
deaths_data <- deaths_data[!(deaths_data$Lat == 0 | is.na(deaths_data$Lat)), ]
deaths_data <- deaths_data[!(deaths_data$Long == 0 | is.na(deaths_data$Long)), ]

# Filter both confirmed and deaths data to only include common countries
common_countries <- intersect(confirmed_data$Country.Region, deaths_data$Country.Region)
confirmed_cases <- confirmed_data[confirmed_data$Country.Region %in% common_countries, c("Country.Region", tail(names(confirmed_data), n = 1))]
deaths <- deaths_data[deaths_data$Country.Region %in% common_countries, c("Country.Region", tail(names(deaths_data), n = 1))]

# Calculate risk scores
risk_scores <- round(100 * deaths[, 2] / confirmed_cases[, 2], 2)

# Create data frame with country names and risk scores
risk_scores_df <- data.frame(Country = confirmed_cases$Country.Region, RiskScore = risk_scores)

# Sort risk scores by ascending order and select first row
lowest_risk <- risk_scores_df[order(risk_scores_df$RiskScore), ][1, ]

# If there are multiple areas with lowest risk, select the one with most confirmations
if (sum(!is.na(lowest_risk$Confirmations)) > 0) {
  lowest_risk <- lowest_risk[which(lowest_risk$Confirmations == max(lowest_risk$Confirmations, na.rm = TRUE)), ]
}

# Display the result
cat("The area with the lowest risk score is", lowest_risk$Country, "with a risk score of", lowest_risk$RiskScore, "and", lowest_risk$Confirmations, "confirmations.")

# Sort risk scores by descending order and select first row
highest_risk <- risk_scores_df[order(-risk_scores_df$RiskScore), ][1, ]

# If there are multiple areas with highest risk, select the one with most confirmations
if (sum(!is.na(highest_risk$Confirmations)) > 0) {
  highest_risk <- highest_risk[which(highest_risk$Confirmations == max(highest_risk$Confirmations, na.rm = TRUE)), ]
}

# Display the result
cat("The area with the highest risk score is", highest_risk$Country, "with a risk score of", highest_risk$RiskScore, "and", highest_risk$Confirmations, "confirmations.")

```

### Objective 5
```{r ob5}
# Load the CSV files for COVID-19 data
confirmed_cases <- read.csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv", stringsAsFactors = FALSE)
deaths <- read.csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv", stringsAsFactors = FALSE)

# Extract the country names from the data frames
countries <- unique(confirmed_cases$Country.Region)

# Initialize empty lists for the sums of confirmed cases and deaths
confirmed_sums <- list()
death_sums <- list()

# Calculate the sums of confirmed cases and deaths for each country
for (country in countries) {
  confirmed_sums[[country]] <- sum(confirmed_cases[confirmed_cases$Country.Region == country, -(1:4)])
  death_sums[[country]] <- sum(deaths[deaths$Country.Region == country, -(1:4)])
}

# Convert the sums into data frames
confirmed_df <- data.frame(Country = names(confirmed_sums), Confirmed = unlist(confirmed_sums))
death_df <- data.frame(Country = names(death_sums), Deaths = unlist(death_sums))

# Sort the data frames by the values of confirmed cases and deaths, respectively
confirmed_df <- confirmed_df[order(-confirmed_df$Confirmed),]
death_df <- death_df[order(-death_df$Deaths),]

# Select the top 5 countries with the most confirmed cases and deaths
top_5_confirmed <- head(confirmed_df, 5)
top_5_deaths <- head(death_df, 5)

# Convert the data frames into tables using the kable function
library(knitr)
kable(top_5_confirmed)
kable(top_5_deaths)


```

#```{bash gitlog}
#git log --pretty=format:"%nSubject: %s%nAuthor: %aN%nDate: %aD%nBody: %b"
#```

Added Objective 1 and 2

commit ee4bfd86cc34401222aabf4f82eb9c489ac97cf0
Author: Manuel Vargas <mvargas5897@student.palomar.edu>
  Date:   Sat Apr 8 13:45:25 2023 -0700

Added Objective 4 and outlines Objectives 1-3

commit c177e309adef63c33db1afa46e6d644d74cecb24
Author: Manuel Vargas <mvargas5897@student.palomar.edu>
  Date:   Sat Apr 8 13:42:44 2023 -0700

Objective 4 completed

commit 894f2eb701e55c5e31e3df59aefb0ce65abdd481
Author: Manuel Vargas <mvargas5897@student.palomar.edu>
  Date:   Sat Apr 8 13:40:48 2023 -0700

Created Objective 4

commit 424c1993596e3940a1700c6018ebf33d0c939e5c
Author: Manuel Vargas <mvargas5897@student.palomar.edu>
  Date:   Sat Apr 8 13:36:31 2023 -0700

Completed Objective 5

commit 781f47924647081dc68cc3384e6c952d545296f5
Author: Manuel Vargas <mvargas5897@student.palomar.edu>
  Date:   Sat Apr 8 13:35:22 2023 -0700

Completed Objective 5

commit 12abe576b9a5443bce20b6f212839f7cf0289ec5
Author: Manuel Vargas <mvargas5897@student.palomar.edu>
  Date:   Sat Apr 8 13:24:27 2023 -0700

Completed Objective 5

commit 7fc1ac718608a13d6236a83b57022ec343909dd5
Author: Manuel Vargas <mvargas5897@student.palomar.edu>
  Date:   Sat Apr 8 08:27:47 2023 -0700

Created Objective 4 R file

commit c14ceb15ff09831b0c86ce64aeac81cfb411cc1c
Author: Manuel Vargas <mvargas5897@student.palomar.edu>
  Date:   Sat Apr 8 08:27:21 2023 -0700

Added Objective 5

commit f399136bf6b63f683e5129ba6617e50cd648fb92
Author: Manuel Vargas <mvargas5897@student.palomar.edu>
  Date:   Sat Apr 8 08:12:01 2023 -0700

Selecting Top 5 Countries

commit 6439c7379e2a90467d25d64d46565daa821405ad
Author: Manuel Vargas <mvargas5897@student.palomar.edu>
  Date:   Sat Apr 8 08:10:11 2023 -0700

Added formatting to Rmd file

commit e8a643ad209123991f92b6a3812f92f810724a5d
Author: Manuel Vargas <mvargas5897@student.palomar.edu>
  Date:   Sat Apr 8 08:02:06 2023 -0700

Created Personal Project Report

commit c3f5d6f5714a53c9bb189a821841064a2e0f6e37
Author: Manuel Vargas <mvargas5897@student.palomar.edu>
  Date:   Sat Apr 8 07:55:08 2023 -0700

Initalized and Calculated Sums of Confirmed Cases and Deaths

commit 6d945ce50e42fc5a5f1b65a15a9341cccfa623f9
Author: Manuel Vargas <mvargas5897@student.palomar.edu>
  Date:   Sat Apr 8 07:51:25 2023 -0700

Extracting Unique Country Names

commit f07a92fe45489d2bf0001a65c28be7e1c9884699
Author: Manuel Vargas <mvargas5897@student.palomar.edu>
  Date:   Sat Apr 8 07:49:48 2023 -0700

Loading CSV files

commit 1d961b81c69c59c366cfff1c13d8f070dfb9d054
Author: Manuel Vargas <mvargas5897@student.palomar.edu>
  Date:   Sat Apr 8 07:48:13 2023 -0700





