---
title: "Project 1 Markdown MV"
author: "Manuel V."
date: "2023-04-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions
The World Health Organization has recently employed a new data science initiative, *CSIT-165*, that uses data science to characterize pandemic diseases. 
*CSIT-165* disseminates data driven analyses to global decision makers.

*CSIT-165* is a conglomerate comprised of two fabricated entities: *Global Health Union (GHU)* and *Private Diagnostic Laboratories (PDL)*. 
Your and your partner's role is to play a data scientist from one of these two entities.


## Data information
Data for 2019 Novel Coronavirus is operated by the John Hopkins University Center for Systems Science and Engineering (JHU CSSE). 
Data includes daily time series CSV summary tables, including confirmations, recoveries, and deaths. Country/region are countries/regions hat conform to World Health Organization (WHO). 
Lat and Long refer to coordinates references for the user. 
Date fields are stored in MM/DD/YYYY format.

## Project Objectives

### Objective 1

```{r ob1}

```

### Objective 2

```{r ob2}

```

### Objective 3

```{r ob3}

```

### Objective 4

```{r ob4}
# Import the data for global confirmed cases
confirmed_url <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv"
confirmed_data <- read.csv(confirmed_url, header = TRUE, stringsAsFactors = FALSE)

# Import the data for global deaths
deaths_url <- "https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv"
deaths_data <- read.csv(deaths_url, header = TRUE, stringsAsFactors = FALSE)

# Remove cruise ships from the confirmed cases data
confirmed_data <- confirmed_data[!(confirmed_data$Lat == 0 | is.na(confirmed_data$Lat)), ]
confirmed_data <- confirmed_data[!(confirmed_data$Long == 0 | is.na(confirmed_data$Long)), ]

# Remove cruise ships from the deaths data
deaths_data <- deaths_data[!(deaths_data$Lat == 0 | is.na(deaths_data$Lat)), ]
deaths_data <- deaths_data[!(deaths_data$Long == 0 | is.na(deaths_data$Long)), ]

# Filter both confirmed and deaths data to only include common countries
common_countries <- intersect(confirmed_data$Country.Region, deaths_data$Country.Region)
confirmed_cases <- confirmed_data[confirmed_data$Country.Region %in% common_countries, c("Country.Region", tail(names(confirmed_data), n = 1))]
deaths <- deaths_data[deaths_data$Country.Region %in% common_countries, c("Country.Region", tail(names(deaths_data), n = 1))]

# Calculate risk scores
risk_scores <- round(100 * deaths[, 2] / confirmed_cases[, 2], 2)

# Create data frame with country names and risk scores
risk_scores_df <- data.frame(Country = confirmed_cases$Country.Region, RiskScore = risk_scores)

# Sort risk scores by ascending order and select first row
lowest_risk <- risk_scores_df[order(risk_scores_df$RiskScore), ][1, ]

# If there are multiple areas with lowest risk, select the one with most confirmations
if (sum(!is.na(lowest_risk$Confirmations)) > 0) {
  lowest_risk <- lowest_risk[which(lowest_risk$Confirmations == max(lowest_risk$Confirmations, na.rm = TRUE)), ]
}

# Display the result
cat("The area with the lowest risk score is", lowest_risk$Country, "with a risk score of", lowest_risk$RiskScore, "and", lowest_risk$Confirmations, "confirmations.")

# Sort risk scores by descending order and select first row
highest_risk <- risk_scores_df[order(-risk_scores_df$RiskScore), ][1, ]

# If there are multiple areas with highest risk, select the one with most confirmations
if (sum(!is.na(highest_risk$Confirmations)) > 0) {
  highest_risk <- highest_risk[which(highest_risk$Confirmations == max(highest_risk$Confirmations, na.rm = TRUE)), ]
}

# Display the result
cat("The area with the highest risk score is", highest_risk$Country, "with a risk score of", highest_risk$RiskScore, "and", highest_risk$Confirmations, "confirmations.")


```

### Objective 5
```{r ob5}
# Load the CSV files for COVID-19 data
confirmed_cases <- read.csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_confirmed_global.csv", stringsAsFactors = FALSE)
deaths <- read.csv("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_time_series/time_series_covid19_deaths_global.csv", stringsAsFactors = FALSE)

# Extract the country names from the data frames
countries <- unique(confirmed_cases$Country.Region)

# Initialize empty lists for the sums of confirmed cases and deaths
confirmed_sums <- list()
death_sums <- list()

# Calculate the sums of confirmed cases and deaths for each country
for (country in countries) {
  confirmed_sums[[country]] <- sum(confirmed_cases[confirmed_cases$Country.Region == country, -(1:4)])
  death_sums[[country]] <- sum(deaths[deaths$Country.Region == country, -(1:4)])
}

# Convert the sums into data frames
confirmed_df <- data.frame(Country = names(confirmed_sums), Confirmed = unlist(confirmed_sums))
death_df <- data.frame(Country = names(death_sums), Deaths = unlist(death_sums))

# Sort the data frames by the values of confirmed cases and deaths, respectively
confirmed_df <- confirmed_df[order(-confirmed_df$Confirmed),]
death_df <- death_df[order(-death_df$Deaths),]

# Select the top 5 countries with the most confirmed cases and deaths
top_5_confirmed <- head(confirmed_df, 5)
top_5_deaths <- head(death_df, 5)

# Convert the data frames into tables using the kable function
library(knitr)
kable(top_5_confirmed)
kable(top_5_deaths)


```
